---
title: "Building Movie Collaboration Network"
author: "by Ta-Chou Ng (Vincent)"
## output: github_document
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, message = F, warning = F)
```

This document demonstrate how to build movie collaboration network, using the credit list scraped from [IMDb](https://www.imdb.com/). We also provide the codes to visualize the network as in the shiny app.

### 1. Build the network  {#sec1}

The annotated code for building the network is as follows.
```{r eval=FALSE}
### movie multiplex network 
require(tidyverse)
require(future)
require(future.apply)
require(igraph)

df1 <- readRDS("dat/Data_info_00to20.RDS")
df2 <- readRDS("dat/Data_people_00to20.RDS")

### filter movies with no overall ratings, outside USA, not English
sel <- which(!is.na(df1$rating_all) & df1$country_USA==1 & df1$lang_English==1 & !is.na(df1$gross_usa)) 
df1 <- df1[sel,] %>%
  select(-starts_with("lang_"),-starts_with("country_"))
df2 <- df2[sel,] %>%
  select(imdb_ID, director, writer, cast)# 6152 movies remaining

### Create edge list (Un-directed, weighted)
### Its empty now, will fill in the edge weights later
any(duplicated(df2$imdb_ID))

N <- NROW(df2$imdb_ID)
edls <- lapply(1:(N-1), function(i){
    tibble(from = i, to = (i+1):N )
})%>% bind_rows()

### Calculate weighted edges between two movies, x specified the type collaboration focused
calc_wts <- function(i, j, x = "cast"){
  #i <- 3533; j <- 5649
  c_ij <- length(intersect(df2[[x]][[i]], df2[[x]][[j]]))
  if(c_ij == 0){ 
    return(c(c_ij = 0, w_ij = 0))
  } else{
    w_ij = c_ij/sqrt(length(df2[[x]][[i]])*length(df2[[x]][[j]]))
    return(c(c_ij = c_ij, w_ij = w_ij))
  }
}

#### Iterated through the edge list, and calculate all three type of edges
plan(multisession, workers = 8)
# co-cast
wts_c <- future_lapply(1:NROW(edls), function(k){
  calc_wts(i = edls$from[k], j = edls$to[k], x = "cast")
}) %>% bind_rows() %>% setNames(nm = c("c_c", "w_c"))
# co-director
wts_d <- future_lapply(1:NROW(edls), function(k){
  calc_wts(i = edls$from[k], j = edls$to[k], x = "director")
}) %>% bind_rows() %>% setNames(nm = c("c_d", "w_d"))
# co-writer
wts_w <- future_lapply(1:NROW(edls), function(k){
  calc_wts(i = edls$from[k], j = edls$to[k], x = "writer")
}) %>% bind_rows()%>% setNames(nm = c("c_w", "w_w"))

plan(sequential)

edls <- bind_cols(edls, wts_c, wts_d, wts_w)
edls <- edls %>%
  filter(c_c + c_d + c_w > 0)

saveRDS(edls, file = "dat/Net_movie_edls.RDS")



### Convert the edge list to an igraph object for calculating node centrality
edls <- readRDS(file = "dat/Net_movie_edls.RDS")%>%
  mutate( from = df1$imdb_ID[from], to = df1$imdb_ID[to])

Gm <- graph_from_edgelist(as.matrix(edls[,c("from","to")]), directed = F)%>%
  set_edge_attr("c_c", value = edls$c_c)%>%
  set_edge_attr("c_d", value = edls$c_d)%>%
  set_edge_attr("c_w", value = edls$c_w)
summary(Gm)

### create the node list
ndls <- as_data_frame(Gm, what = "vertices") %>%
  left_join(df1, by = c("name"="imdb_ID")) %>%
  rename(id = name)

### calculate node eigen centrality in terms of all types of edges
tmp <- eigen_centrality(
  Gm %>% set_edge_attr("weight", value = edls$w_c))$vector
ndls$eigen_cen_c <- tmp[ndls$id]

tmp <- eigen_centrality(
  Gm %>% set_edge_attr("weight", value = edls$w_d))$vector
ndls$eigen_cen_d <- tmp[ndls$id]

tmp <- eigen_centrality(
  Gm %>% set_edge_attr("weight", value = edls$w_w))$vector
ndls$eigen_cen_w <- tmp[ndls$id]


```

### 2. Visualize the network  {#sec2}

Then we visualize the network interactively using the `visNetwork` R package.

```{r}
require(tidyverse)
require(visNetwork)

# load the calculated network
load(file = "../dat/Net_movie.RData")

ndls0 <- left_join(ndls, df2, by = c("id" = "imdb_ID"))%>%
  select(-starts_with("genre_"), -ends_with("male"), -ends_with("female"))%>%
  mutate(
    director = lapply(director, function(x) paste0(x, collapse = ", ") ) %>% unlist,
    writer = lapply(writer, function(x) paste0(x, collapse = ", ") )%>% unlist,
    cast = lapply(cast, function(x) paste0(x[1:5], collapse = ", ") )%>% unlist,
    genres = lapply(genres, function(x) paste0(x[1:3], collapse = ", ") )%>% unlist,
    award_win = ifelse(is.na(award_win), 0, award_win),
    title = str_remove(title, "\\(.+\\)"),
  )%>%
  filter(title != "")%>%
  arrange(title)

# helpers
mvn2id <- setNames(ndls0$id, nm = ndls0$title)
id2mvn <- setNames(ndls0$title, nm = ndls0$id)
id2rating <- setNames(ndls0$rating_all, nm = ndls0$id)

# define tooltip info
ndls <- ndls0 %>%
  mutate(
    tooltip = paste0("<div style='font-size:70%'>",
                   "<b style='font-size:150%;color:#396a93;'>", title,"</b> <b>(", year, ")</b><br>",
                   "<b>Director:</b> ", director, "<br>",
                   "<b>Writer:</b> ", writer, "<br>",
                   "<b>Cast:</b> ", cast, ", ...<br>",
                   "<b>Genre:</b> ", genres, "<br>",
                   "<b>Rating:</b> ", rating_all, "/10 by ", vote_all, " voters <br>",
                   "<b>Box office:</b> ", gross_usa, " US dollars<br>",
                   "<b>Awards winned:</b> ", award_win, "<br>",
                   "<b>Eigen centrality:</b> ", round(eigen_cen_c, 3),"<br>",
                   "<b>More on <a href='https://www.imdb.com/title/", id ,"'>IMDb</a></b></div>"
                   )
  ) %>%
  select(id, title, year,award_win, rating_all, vote_all, gross_usa, cumulative_worldwide_gross, budget, tooltip)%>%
  mutate(
    rating_all_vote = rating_all*vote_all,
    rating_all_vote = ecdf(rating_all_vote)(rating_all_vote),
    rating_all  = ecdf(rating_all)(rating_all),
    gross_usa  = ecdf(gross_usa)(gross_usa),
    cumulative_worldwide_gross  = ecdf(cumulative_worldwide_gross)(cumulative_worldwide_gross),
    award_win = ecdf(award_win)(award_win)
    )


edls <- edls %>%
  mutate(title = paste0("<div style='font-size:70%'>",
                        "<b style='color:#396a93;'>", id2mvn[from], " <-> ",
                        id2mvn[to], "</b><br>",
                        "<b>No. of co-star:</b> ", c_c, "<br>",
                        "<b>No. of co-director:</b> ", c_d, "<br>",
                        "<b>No. of co-writer:</b> ", c_w, "</div>" ))


# Main function to plot the ego network of a movie
plot_mv_ego <- function(mvn, evar = "c_d",
                        max.n = 100, nodeLS = ndls,  edgeLS = edls, 
                        m2i = mvn2id, i2r = id2rating, yvar = "rating_all"){

  x <- m2i[[mvn]]
  ed_x <- edgeLS %>%
      filter(from == x | to == x)%>%
      filter(!!sym(evar) > 0)%>%
      mutate(color = "#a6a6a6")
  
  x.nei <- unique(c(ed_x$from, ed_x$to))%>% .[.!=x]    
  
  wcut <- ifelse(evar == "c_c", 3, 1)
  ed_x <- bind_rows(
    ed_x,
    edgeLS %>% filter(!!sym(evar) > wcut)%>%
      filter(from %in% x.nei & to %in% x.nei)%>%
      mutate(color = "#cccccc")
  ) 
  
  if(length(x.nei) > max.n){ # show at most max.n nodes
    x.nei <- i2r[x.nei] %>% sort(decreasing = T) %>%
      names() %>% .[1:max.n]
  }
  
  nd_x <- nodeLS %>%
    filter(id %in% c(x, x.nei))%>%
    mutate(color = ifelse(id == x, "#DC143C", "#7eaacd"),
           label = "", value = !!sym(yvar), title = tooltip) 
  
  out <- visNetwork(nd_x, ed_x, background ="#fff9f3")%>% 
    visEdges(
       # color = list(hover="#eb2d53",
       #            highlight="#eb2d53", inherit=F),
      scaling = list(min=.5, max=25))%>%
    visNodes(borderWidth=0.3,
             scaling = list(min=5, max=25))%>%
    visOptions(
      highlightNearest =list(enabled=T, 
                             algorithm="hierarchical", degree=1))%>%
    visInteraction(navigationButtons = F, hover = T,
                   hideEdgesOnDrag=T, selectConnectedEdges=T)%>%
    visPhysics(enabled = T,
               barnesHut = list(
                 centralGravity = 0.3,
                 springConstant = 0.04,
                 damping = 1
               ))
  return(out)
}
```


Then we test the plotting function using Monsters University as an example

```{r}
plot_mv_ego(mvn = "Monsters University", evar = "c_c", max.n = 50, yvar = "gross_usa")
```



Then, we packed all the data, functions for the siny app module.
```{r eval=FALSE}
save(ndls, edls, plot_mv_ego, mvn2id, id2rating, file = "dat/Module2_network.RData")
```

